---
title: "Fall 2022 Galleria field infection data"
author: "Morgan Swoboda"
date: "2023-02-08"
output: github_document
---

# Package install and upload data sheet
```{r, eval = TRUE}

#install.packages("emmeans")
#install.packages("ggthemes")

library(ggplot2)
library(emmeans)
library(tidyr)
library(dplyr)
library(ggthemes)
library(RColorBrewer)

#import data
Fall22EPF <- read.csv("Fall 22 epf bioassay - Sheet1.csv", )

#make treatments and varieties into factors
Fall22EPF$treatment = factor(Fall22EPF$treatment, levels = c("Control", "10Met", "10Bb"))
Fall22EPF$variety = factor(Fall22EPF$variety, levels = c("Cochise", "Armani"))
#Fall22EPF$num.dead = as.numeric(Fall22EPF$num.dead)
#Fall22EPF$num.pupae = as.numeric(Fall22EPF$num.pupae)
#Fall22EPF$total.out = as.numeric(Fall22EPF$total.out)
#Fall22EPF$num.met.infect = as.numeric(Fall22EPF$num.met.infect)
```

#Prelim data check
```{r, eval=TRUE}
ggplot(data = Fall22EPF, aes(x = treatment, y = perc.inf.total, color = variety)) + geom_boxplot() + ggtitle("Total infection")
#everything looks ok

ggplot(data = Fall22EPF, aes(x = treatment, y = perc.met.total, color = variety)) + geom_boxplot() + ggtitle("Metarhizium infection")
```
# Stats for Metarhizium - normal ANOVA way?
```{r,eval=TRUE}
attach(Fall22EPF)

##two way ANOVA
two.way.Met <- aov(perc.met.total~treatment*variety)
summary(two.way.Met) #no signifcant differences, no interaction, change to one way model instead
summary.lm(two.way.Met)

##one way anova
one.way.Met <- aov(perc.met.total~treatment)
summary(one.way.Met) #not significant
summary.lm(one.way.Met)
plot(one.way.Met) #probably not normal based on QQ plot, looks like it's exponential?
shapiro.test(resid(one.way.Met)) #residuals NOT normal
fligner.test(perc.met.total~treatment) #normal

# KW instead since it's not normal
##KW for met
MetF22KW <- kruskal.test(perc.met.total~treatment)
print(MetF22KW) #not significant

```

# Try a different type of analysis? Could this data be proportions?
## X is categorical, unsure what to do with Y? 
```{r, eval=TRUE}

head(Fall22EPF)
par(mfrow = c(1,2))

inf.proportion <- num.met.infect/total.out #looking for the proportion of metarhizium infections out of the number of removed insects
plot(treatment, inf.proportion, ylab = "proportion met infected", xlab("Treatment"))
plot(treatment, log(inf.proportion), ylab = "(log)proportion met infected", xlab("Treatment")) #messes up the plots because of an outlier?

```

#Try a GLM?
## can use a GLM when the variance is not constant, and/or when the errors are not normally distributed
### Might consider using GLMs when the response variable is: count data expressed as proportions, count data that are not proportions, binary response variables, data on time to death where the varience increases faster than linearly with the mean
```{r, eval=TRUE}
names(Fall22EPF) #num.dead is the number of galleria counted as dead, num.pupae is the number that pupated, total.out is the sum of num.dead and num.pupae, num.met.infect is the number of galleria that exhibited Metarhizium infections

#look at main effect means
tapply(num.met.infect, treatment, mean)
tapply(num.met.infect, variety, mean)

#glm model of interactions between treatment and variety
glmmodel <- glm(num.met.infect ~ treatment * variety, poisson)
summary(glmmodel) #no interaction, remove variety as a main effect

#R reports two forms of deviance – the null deviance and the residual deviance. The null deviance shows how well the response variable is predicted by a model that includes only the intercept (grand mean). It’s a measure of badness of fit–higher numbers indicate worse fit.

#The Akaike Information Criterion (AIC) provides a method for assessing the quality of your model through comparison of related models.  It’s based on the Deviance, but penalizes you for making the model more complicated.  Much like adjusted R-squared, it’s intent is to prevent you from including irrelevant predictors. However, unlike adjusted R-squared, the number itself is not meaningful. If you have more than one similar candidate models (where all of the variables of the simpler model occur in the more complex models), then you should select the model that has the smallest AIC. So it’s useful for comparing models, but isn’t interpretable on its own.

model2 <-  glm(num.met.infect ~ treatment, poisson)
summary(model2)

#see how the models compare to each other using an anova
anova(glmmodel,model2, test = "Chi") #not significantly different, so we're ok using the model2
plot(model2)

#test the model that you think fits best with the Hosmer Lemeshow goodness of fit test.
# install.packages("ResourceSelection") #need this package for the hoslem.test() function.
library(ResourceSelection)

hoslem.test(Fall22EPF$num.met.infect, fitted(model2))
```
# Plot the GLM model
```{r, eval=TRUE}

```

